Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-05-09 exp1/iswc2013.tex
make github a \url
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@542 43ef1190-90fd-4dd7-a769-d0121f02b80e
data factors todo on 32 run
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@543 43ef1190-90fd-4dd7-a769-d0121f02b80e

All the code to reproduce this study as well as detailed instructions
is published on Github: \texttt{https://github.com/kjetilk/doe-sparql}.

@@

The resulting normal plot is in figure~\ref{fig:frac32normal}. We see that
only two main effects are significant on a $\alpha = 0.05$ level
according to the Lenth criterion, ``Implement'' is again deemed a
significant improvement, and ``Union'' is deemed significantly the
hardest operation for the query engine. We see that some other points
also deviate from the straight line, but with as few runs as this, the
total variance is great, so no other conclusions can be drawn. Also,
the Student's t-test we used in the previous section cannot be
employed, as we have no additional data to use for the test.

@@

The present experiment is very primitive in its choice of factors, and
we believe that the strategy employed to construct suitable
experiments will be the main determinant for the long-term feasibility
of this direction. One strategy will be to start with broad factors
such as ``Machine'' and refine them with experience. A more difficult
issue is how to address the problem of testing the whole SPARQL
language\cite{sparql11query}. We believe that the parameterization work inititated by
\cite{goerlitz2012splodge} with their SPLODGE system is a key
ingredient. They parameterized SPARQL queries in such a way they could
be autogenerated. One suggestion is to parameterize queries based on
the grammar, so that one factor becomes the number of e.g. UNIONs, and
the levels are chosen. One problem one will soon run into is that
parts of the language is recursive, e.g. a GroupGraphPattern can
consist of several GroupGraphPatterns. However, we believe that
pragmatic limits can be set, it is for example not of practical
interest to nest OPTIONALs to any great depth even if it is allowed by
the language.

>>>>

All the code to reproduce this study as well as detailed instructions
is published on Github: \url{https://github.com/kjetilk/doe-sparql}.

@@

The resulting normal plot is in figure~\ref{fig:frac32normal}. We see that
only two main effects are significant on a $\alpha = 0.05$ level
according to the Lenth criterion, ``Implement'' is again deemed a
significant improvement, and ``Union'' is deemed significantly the
hardest operation for the query engine. We see that some other points
also deviate from the straight line, but with as few runs as this, the
total variance is great, so no other conclusions can be drawn. Also,
the Student's t-test we used in the previous section cannot be
employed, as we have no additional data to use for the
test.\todo{actually, we can}

@@

The present experiment is very primitive in its choice of factors, and
we believe that the strategy employed to construct suitable
experiments will be the main determinant for the long-term feasibility
of this direction. One strategy will be to start with broad factors
such as ``Machine'' and refine them with experience. Also, the data
must be parameterized with many more factors than just the number of
triples; data heterogenity \cite{Duan:2011:AOC:1989323.1989340} is one
example, but also queries over highly skewed data is important. A more
difficult issue is how to address the problem of testing the whole
SPARQL language\cite{sparql11query}. We believe that the
parameterization work inititated by \cite{goerlitz2012splodge} with
their SPLODGE system is a key ingredient. They parameterized SPARQL
queries in such a way they could be autogenerated. One suggestion is
to parameterize queries based on the grammar, so that one factor
becomes the number of e.g. UNIONs, and the levels are chosen. One
problem one will soon run into is that parts of the language is
recursive, e.g. a GroupGraphPattern can consist of several
GroupGraphPatterns. However, we believe that pragmatic limits can be
set, it is for example not of practical interest to nest OPTIONALs to
any great depth even if it is allowed by the language.
